Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/splitting.tex
Half-baked. Just checking in to merge.

In this section, we present a \emph{splitting algorithm} which
statically converts a term in \lang\ into a pair of separate,
monostaged terms that individually correspond to the first and second
stage computations of the \lang\ term. Due to important differences in
structure, we describe both the correctness condition and algorithm
for splitting of \bbonem\ terms separately from that of \bbtwo\ terms.
By construction, \bbonep\ terms are monostage and do not require
splitting.

@@

%More precisely
For any $\coltwo e A$ which reduces to a residual $q$ (via $\diatwo e q$),
$e$ splits into a syntactically separated program $\mathcal{P}$ 
(via the new relation $e \splittwosym \mathcal{P}$) which is
reduced to a residual $q'$ (via the new relation $\mathcal{P} \sepredtwosym q'$), which is the same as $q$ up to an equivalence.
By ``syntactically separated'' we mean that $\mathcal{P}$ 
has the form $\pipeS p l r$, where  
$p$ (called the {\em precomputation}) is a monostage term representing all the
first-stage computations in $e$, 
and $l.r$ (called the {\em resumer}) is a monostage term representing all the
second-stage computations in $e$.  
$\sepredtwosym$ is defined by a single rule, which evaluates the precomputation and substitutes the resulting value in for $l$ in the resumer:
\[
\infer{\sepredtwo {\pipeS p l r} {[b/l]r}} {\reduce p b}
\]
We call $b$ the {\em boundary value} resulting from the split since it represents the data communicated between the first and second stage computations.
A summary of these operations and the relationship between them is provided on the right side of \ref{fig:splittingSummary}.

>>>>

In this section, we present a \emph{splitting algorithm} which
statically converts a term in \lang\ into a pair of separate,
monostaged terms that individually correspond to the first and second
stage computations of the \lang\ term. Due to important differences in
structure, we describe both the splitting correctness condition and
splitting algorithm for \bbonem\ terms separately from that of
\bbtwo\ terms.  By construction, \bbonep\ terms are monostage and do
not require splitting.

@@

%More precisely

As described in \ref{sec:example},

For any $\coltwo e A$ which reduces to a residual $q$ (via $\diatwo e q$),
$e$ splits into a syntactically separated program $\mathcal{P}$ 
(via the new relation $e \splittwosym \mathcal{P}$) which is
reduced to a residual $q'$ (via the new relation $\mathcal{P} \sepredtwosym q'$), which is the same as $q$ up to an equivalence.
By ``syntactically separated'' we mean that $\mathcal{P}$ 
has the form $\pipeS p l r$, where  
$p$ (called the {\em precomputation}) is a monostage term representing all the
first-stage computations in $e$, 
and $l.r$ (called the {\em resumer}) is a monostage term representing all the
second-stage computations in $e$.  
$\sepredtwosym$ is defined by a single rule, which evaluates the precomputation and substitutes the resulting value in for $l$ in the resumer:
\[
\infer{\sepredtwo {\pipeS p l r} {[b/l]r}} {\reduce p b}
\]
We call $b$ the {\em boundary value} resulting from the split since it represents the data communicated between the first and second stage computations.
A summary of these operations and the relationship between them is provided on the right side of \ref{fig:splittingSummary}.
